Regression

### 2015-08-26 cluster.rmd
update randoom forest

>>>>

Adjust inv.mlsto, cost.mlsto, balance.mlsto, for inflation!!!!

library('quantmod', lib= 'C:/Progra~1/R/R-3.2.2/library')
getSymbols("CPIAUCSL", src='FRED') #Consumer Price Index for All Urban Consumers: All Items
#[1] "CPIAUCSL"
tail(CPIAUCSL)

# make an df object of costs
p<- data.frame('Year' = all7d$Year, 'inv.mlsto'= all7d$inv.mlsto, 'balance.mlsto' = all7d$balance.mlsto,
               'cost.mlsto' = all7d$cost.mlsto, 'mlsto' = all7d$mlsto)

#calculate annual averave inflation since 2004
avg.cpi <- apply.yearly(CPIAUCSL, mean)
#ratio using 2014 as base year
cf <- avg.cpi/as.numeric(avg.cpi['2014'])
#create new cf column for date and extract year
cf<- as.data.frame(cf)
cf$Date<- rownames(cf)
cf$Date<- cf$Date %>% as.Date()
cf$Date<- format(cf$Date,"%Y")
p<- merge(p, cf, by.x='Year', by.y= 'Date')
#divide all dollar values by CPIAUCSL column
p<- transform(p, adj.inv.mlsto = inv.mlsto/CPIAUCSL,
          adj.balance.mlsto = balance.mlsto/CPIAUCSL,
          adj.cost.mlsto = cost.mlsto/CPIAUCSL)
#check
s<- sample(1:nrow(all7d),10)
p[s,]
#replace existing all7d columns
all7d<- merge(all7d, p %>% select(mlsto, adj.inv.mlsto, adj.balance.mlsto, adj.cost.mlsto) , by='mlsto')
all7d<- all7d %>% select(-inv.mlsto, - balance.mlsto, -cost.mlsto)
colnames(all7d)[names(all7d) %in% "adj.inv.mlsto"] <- "inv.mlsto"
colnames(all7d)[names(all7d) %in% "adj.balance.mlsto"] <- "balance.mlsto"
colnames(all7d)[names(all7d) %in% "adj.cost.mlsto"] <- "cost.mlsto"

write.csv(all7d, 'all7d.csv')

### 2015-08-26 rand_forest.rmd
update randoom forest

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library('caret',lib='C:/Progra~1/R/R-3.2.1/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library('fpc', lib='C:/Progra~1/R/R-3.2.1/library')
# library('vegan', lib = 'C:/Progra~1/R/R-3.2.1/library')
# library(devtools,lib='C:/Progra~1/R/R-3.2.1/library')

@@

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
all7d$Post.Code<- as.factor(all7d$Post.Code)

@@

#create tree with training data only
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
summary(tree.all7d)

#have a look at cp values, automatic cross validation in rpart
printcp(tree.all7d)
plotcp(tree.all7d)
#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.011))

@@

prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
#make prediction with current tree
tree.pred<-predict(tree.all7d, test)
#calculate RMSE for test data
RMSE(test$return.pdol, tree.pred)
# 51.4 cents, already close to regression..

@@

RMSE is 0.514 which is worse than regression in predicting return per dollar.

@@

#RMSE = 0.50386
# Rsquared = 0.0844, quite bad

@@

@@

        #perform 10 fold xval using caret package
        tc = trainControl("cv",10)
        #keep cp static at 0.011
        rpart.grid = expand.grid(.cp=0.011)
        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

@@

addone(var='JD.Second', df=all7d)
#RMSE = .4709
addone(var='Post.Code', df=all7d)
#RMSE = .4353
addone(var='majority.pos', df=all7d)
#RMSE = .4541
addone(var='pc.majpos.log', df=all7d)
#RMSE = .4553
addone(var='code.client', df=all7d)
#RMSE = .4394
addone(var='Billing.Type', df=all7d)
#RMSE = .4579
addone(var='code.contact', df=all7d)
#RMSE = .3972

#dont use post code ever. 
#JD.Second doesn't help that much. acutally makes RMSE worse.

@@

Delete Post.Code option

Billing.Type should be paired with inv.mlsto and maybe Discipline

maybe client should be paired with inv.mlsto
maybe JD.Second should be paired with inv.mlsto and code.client and timespan and pc.pro and no.users?? which ever one does something.

majority.pos pair with inv.mlsto
pc.majpos pair with inv.mlsto

Write a function that lists the extra variable you're trying and write out the interaction.

@@

Now look at cross validation for full variable set:

#using caret package
tc <- trainControl("cv",10,savePred=F)
fit<- train(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
              data=all7d,
            method = 'lm', trControl = tc)

#RMSE = 0.49246
#compared to non-cross validated
# RMSE = 0.4759
```

Not really trying to tune anything via cross validation, however RMSE increased from 0.4759 to 0.49246.

Not THAT bad.

>>>>

library('knitr', lib = 'C:/Progra~1/R/R-3.2.2/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.2/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.2/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.2/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.2/library')
library('caret',lib='C:/Progra~1/R/R-3.2.2/library')
library('randomForest',lib='C:/Progra~1/R/R-3.2.2/library')
library('rpart',lib='C:/Progra~1/R/R-3.2.2/library')
library('rpart.plot',lib='C:/Progra~1/R/R-3.2.2/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.2/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('randomForest', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library('fpc', lib='C:/Progra~1/R/R-3.2.2/library')
# library('vegan', lib = 'C:/Progra~1/R/R-3.2.2/library')
# library(devtools,lib='C:/Progra~1/R/R-3.2.2/library')

@@

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
all7d$Post.Code<- as.factor(all7d$Post.Code)

@@

#create tree with training data only
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
# summary(tree.all7d)

#have a look at cp values, automatic cross validation in rpart
printcp(tree.all7d)
plotcp(tree.all7d)
#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train,
                  control = rpart.control(cp= 0.013))

@@

prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
#make prediction with current tree
tree.pred<-predict(tree.all7d, test)
#calculate RMSE for test data
RMSE(test$return.pdol, tree.pred)
# 49.8 cents, already close to regression..

@@

RMSE is 0.498 which is worse than regression in predicting return per dollar.

@@

#RMSE = 0.505
# Rsquared = 0.095, quite bad

@@

        #select cp
        cp.new = add.one$cptable[which.min(add.one$cptable[,"xerror"]),"CP"]-0.0002
        add.one = rpart(as.formula(formula), data = reduced,
                        control = rpart.control(cp= cp.new))

@@

        #perform 10 fold xval using caret package
        tc = trainControl("cv",10)
        #keep cp static at 0.011
        rpart.grid = expand.grid(.cp=cp.new)
        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

@@

addone(var='JD.Second', df=all7d)
#RMSE = .528
addone(var='Post.Code', df=all7d)
#RMSE = .499
addone(var='majority.pos', df=all7d)
#RMSE = .49
addone(var='pc.majpos', df=all7d)
#RMSE = .486
addone(var='code.client', df=all7d)
#RMSE = .534
addone(var='Billing.Type', df=all7d)
#RMSE = .502
addone(var='code.contact', df=all7d)
#RMSE = .535

# none of the extra variables actually appear in the revised trees except JD.Second
# can only assume the different subsets are what is causing the variation in RMSE

@@

Write a function that can plot a tree
Limit Year
choose additional variables to include.

@@
